### 2021-06-16 thesis/tex/ICtheory.tex
update

\subsection*{Monte Carlo normalization}
Independent researchers outside of the IceCube collaboration will not be able to more persicely
simulate the detector. The IceCube Monte Carlo is a complex and proprietary machinery, so our goal in this 
section is to come as close as we can to their Monte Carlo simulations. After we are confident that 
our code displays the same overall features as the `offical', we normalize our results $N_{ij}^\text{sim}$ as 
\begin{align}\label{eq:MC_norm}
    N_{ij} = \frac{N_{ij}^\text{null}}{N_{ij}^\text{MC}} N_{ij}^\text{sim}\,.
\end{align}
For each bin $i,j$, we then obtain a correction factor which contains information that we are unable
to obtain or sufficiently incorporate. One example of such information is the systematic errors of the DOMs.
Recent IceCube data releases do not include such information. Since the systematic errors are affecting the 
event count on a bin-by-bin basis, they can in theory drastically modify the binned results. Another example of
an error source what will be remedied by this method is the flux. We are using a fairly simple model of the atmospheric 
flux that excludes prompt flux. The IceCube collaboration use several different flux models which are initialized 
by a parametrization of the cosmic ray flux.\footnote{Included in the cosmic ray models are e.g. the pion to kaon 
ratio, which are often used as a nuisance parameter. By not being able to include this in our error analysis, our method will 
be limited to only consider the overall flux normalization, rather than the components that produce the flux in the first place.}
%TODO: maybe rambling?

In Fig.~\ref{fig:IC_MC_norm}, we present the Icecube Monte Carlo obtained from their 2020 sterile analysis~\cite{IC2020}, along
with our null hypothesis times a constant factor.  We deemed these shapes to be satisfactory, thus allowing us to multiply Eq.~\ref{eq:ICevents} by the 
correction factors of Eq.~\ref{eq:MC_norm}. We now arrive to our final event count
\begin{align}\label{eq:Nth}
    N_{ij} &= \frac{N_{ij}^\text{null}}{N_{ij}^\text{MC}} T \sum_\beta \int_{(\cos{\theta_z^r})_i}^{(\cos{\theta_z^r})_{i+1}} \dd \cos{\theta^r_z} \int_{E^r_{j}}^{E^r_{j+1}} \dd E^r 
    \int_{E^t_{min}}^{E^t_{max}} R(E^r,E^t) \phi_\beta^\text{det}  A^\text{eff}_\beta\, \dd E^t
    \,,
\end{align}
with $E^t_{min}$, $E^t_{max}$, and $R(E^r,E^t)$ taken from the Gaussian process regressor.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/IC_MC_norm.pdf}
    \caption{IceCube Monte Carlo, binned in $\Ereco$ and $\zreco$. We compare this with our simulations shown as ´Null' in the plots.}\label{fig:IC_MC_norm}
\end{figure} 
%TODO:fill more
% \newpage
% \bibliographystyle{nature}
% \bibliography{ref.bib}
% \end{document}

>>>>

\subsection*{Monte Carlo normalization}
Independent researchers outside of the IceCube collaboration will not be able to more persicely
simulate the detector. The IceCube Monte Carlo is a complex and proprietary machinery, so our goal in this 
section is to come as close as we can to their Monte Carlo simulations. After we are confident that 
our code displays the same overall features as the `offical', we normalize our results $N_{ij}^\text{sim}$ as 
\begin{align}\label{eq:MC_norm}
    N_{ij} = \frac{N_{ij}^\text{null}}{N_{ij}^\text{MC}} N_{ij}^\text{sim}\,.
\end{align}
For each bin $i,j$, we then obtain a correction factor which contains information that we are unable
to obtain or sufficiently incorporate. One example of such information is the systematic errors of the DOMs.
Recent IceCube data releases do not include such information. Since the systematic errors are affecting the 
event count on a bin-by-bin basis, they can in theory drastically modify the binned results. Another example of
an error source what will be remedied by this method is the flux. We are using a fairly simple model of the atmospheric 
flux that excludes atmospheric prompt and astrophysical fluxes. The IceCube collaboration use several different flux models which are initialized 
by a parametrization of the cosmic ray flux.\footnote{Included in the cosmic ray models are e.g. the pion to kaon 
ratio, which are often used as a nuisance parameter. By not being able to include this in our error analysis, our method will 
be limited to only consider the overall flux normalization, rather than the components that produce the flux in the first place.}

In Fig.~\ref{fig:IC_MC_norm}, we present the Icecube Monte Carlo obtained from their 2020 sterile analysis~\cite{IC2020}, along
with our null hypothesis times a constant factor.  We deemed these shapes to be satisfactory, thus allowing us to multiply Eq.~\ref{eq:ICevents} by the 
correction factors of Eq.~\ref{eq:MC_norm}. We now arrive to our final event count
\begin{align}\label{eq:Nth}
    N_{ij} &= \frac{N_{ij}^\text{null}}{N_{ij}^\text{MC}} T \sum_\beta \int_{(\cos{\theta_z^r})_i}^{(\cos{\theta_z^r})_{i+1}} \dd \cos{\theta^r_z} \int_{E^r_{j}}^{E^r_{j+1}} \dd E^r 
    \int_{E^t_{min}}^{E^t_{max}} R(E^r,E^t) \phi_\beta^\text{det}  A^\text{eff}_\beta\, \dd E^t
    \,,
\end{align}
with $E^t_{min}$, $E^t_{max}$, and $R(E^r,E^t)$ taken from the Gaussian process regressor.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/IC_MC_norm.pdf}
    \caption{IceCube Monte Carlo, binned in $\Ereco$ and $\zreco$. We compare this with our simulations shown as ´Null' in the plots.}\label{fig:IC_MC_norm}
\end{figure} 
Thus, we are now able to sufficiently approximate the IceCube Monte Carlo, which makes us able to run simulations based on different physics scenarios.
% \newpage
% \bibliographystyle{nature}
% \bibliography{ref.bib}
% \end{document}

### 2021-06-16 thesis/tex/introduction.tex
update

\chapter{Introduction}
There is no reason for us to believe that our current knowledge in particle physics is complete.
The most successful framework -- the Standard Model -- has notable shortcomings. Neither does it 
explain why there is something rather than nothing, nor does it give a candidate for dark matter\footnote{In the Standard Model,
matter and antimatter come in pairs of one particle and one antiparticle. When interacting,
antimatter \emph{annihilates} the matter, resulting in pure light. Since the universe exists,
there must have been more matter than antimatter at the Big Bang. %TODO: check and make better
Dark matter is heavy matter which does not interact
with electromagnetic radiation, and is thus ´invisible' to us. 
No such particle exists in the Standard Model.}.

To make matters even worse, three of the massless particles in the Standard Model do have mass.
In this thesis we focus on the said particles, the neutrinos, and propose extensions to the Standard Model
related to them. We then compare the effects of these extensions with collected data to see 
if our new theory more closely describes Nature or not.

Ultimately, the purpose of this excursion is to guide future research where to uncover more accurate theories.
Traces of this grander theory will be present as breadcrumbs in Nature, we just need to know where 
and how carefully to look.

\section{Outline}
This thesis is outlined as follows. In Chapter~\ref{ch:osc}, we briefly review the most relevant parts 
of the Standard Model that will be relevant for our extensions of it. Then we present the evidence and 
subsequent discovery of neutrino oscillations, which ultimately led to the 2015 Nobel physics prize, 
along with the modifications to the Standard Model needed to accomodate the discovery. The 
reader is then introduced to a standard but detailed derivation of the neutrino mixing matrix
and oscillation Hamiltonian, including the matter 
effects stemming from charged and neutral current weak interactions within the Earth.

In Chapter~\ref{ch:ic}, we present an experimental procedure used to observe signals from neutrinos originating
from cosmic ray interactions in the atmosphere of the Earth. Furthermore, we introduce the reader to 
the two present Antarctic neutrino detectors: IceCube and DeepCore. We also give a brief summary 
about a proposed detector upgrade: Precision IceCube Next Generation Upgrade (PINGU), which 
we will use as a forecast in our analysis.

Chapter~\ref{ch:theory} consists of two parts. In Section~\ref{sec:anomalies}, we 
introduce a new particle: the sterile neutrino. We present how neutrino oscillations are
modified by this hypothesized fourth neutrino, and how these modifications would appear in IceCube if 
the particle is present in Nature. In Section~\ref{sec:nsiTheory}, we completely set the sterile neutrino aside,
and instead turn to the interactions between neutrinos and the Earth. We again amend the Standard Model by 
considering exotic interactions from a higher energy theory, manifesting themselves as sub-leading modifications 
to the matter potential. These new interactions are not present in the Standard Model and are thus called 
non-standard interactions (NSI).

>>>>

\chapter{Introduction}
There is no reason for us to believe that our current knowledge in particle physics is complete.
The most successful framework -- the Standard Model -- has notable shortcomings. Neither does it 
explain why there is something rather than nothing, nor does it give a candidate for dark matter\footnote{In the Standard Model,
matter and antimatter come in pairs of one particle and one antiparticle. When interacting,
antimatter \emph{annihilates} the matter, resulting in pure light. Since the universe exists,
there must have been more matter than antimatter at the Big Bang. %TODO: check and make better
Dark matter is heavy matter which does not interact with electromagnetic radiation.
No such particle exists in the Standard Model.}.

To make matters even worse, three of the massive particles are massless in the Standard Model.
In this thesis, we focus on the said particles -- the neutrinos -- and propose extensions to the Standard Model related to them. 
We then compare the effects of these extensions with collected data to see if our new theory more closely describes Nature or not.

Ultimately, the purpose of this excursion is to guide future research where to uncover more accurate theories.
Traces of this grander theory will be present as breadcrumbs scattered in Nature. We just need to know where and how carefully to look.

\section{Outline}
This thesis is outlined as follows. In Chapter~\ref{ch:osc}, we briefly review the most relevant parts of the Standard Model that will be relevant for our extensions of it.
Then we present the evidence and subsequent discovery of neutrino oscillations, which ultimately led to the 2015 Nobel physics prize, along with the modifications to the Standard Model needed to accommodate the discovery.
The reader is then introduced to a standard but detailed derivation of the neutrino mixing matrix and oscillation Hamiltonian, including the matter effects stemming from charged and neutral current weak interactions within the Earth.

In Chapter~\ref{ch:ic}, we present an experimental procedure used to observe signals from neutrinos originating from cosmic ray interactions in the atmosphere of the Earth. 
Furthermore, present the two present Antarctic neutrino detectors: IceCube and DeepCore.
We also summarize a proposed detector upgrade: Precision IceCube Next Generation Upgrade (PINGU), which we will use as a forecast in our analysis.

Chapter~\ref{ch:theory} consists of two parts. In Section~\ref{sec:anomalies}, we introduce a new particle: the sterile neutrino. 
We present how neutrino oscillations are modified by this hypothesized fourth neutrino and how these modifications would appear in IceCube if the particle is present in Nature. 
In Section~\ref{sec:nsiTheory}, we completely set the sterile neutrino aside and instead turn to the interactions between neutrinos and the Earth. 
We again amend the Standard Model by considering exotic interactions from a higher energy theory, manifesting themselves as sub-leading modifications to the matter potential. 
Since these new interactions are not present in the Standard Model, we call them non-standard interactions (NSI).

### 2021-06-16 thesis/tex/nsiTheory.tex
update

Compared to other fermions, the lightness of neutrino masses along with their sparse SM interactions might 
indicate that these particles provide the best starting point for us to probe new physics, in which 
other neutrino interactions might occur~\cite{gavela2009}. We call these interactions non-standard interactions (NSI).

Following the approach of the Standard Model that the group generators uniquely determine the gauge bosons, a different gauge
theory will have different interactions than those we presently know. These new interactions can be parametrized as model-independent four-fermion effective operators \cite{salvadoNSI,nsiFarzan}.
Following the discussion in~\cite{tommyNSI}, the six-dimensional NSI parameters $\epsilon$ have the scale
\begin{align}
    \epsilon \propto \frac{m_W^2}{m_{\epsilon}^2} \sim \frac{10^{-2}}{m_\epsilon^2}\,
\end{align} 
in TeV, so the new interactions generated at a mass scale of $m_\epsilon = \SI{1}{TeV}$ will produce parameters in the order of $10^{-2}$, 
two magnitudes below the standard matter effect. Thus, if we assume the new interactions to arise from a higher-energy theory above electroweak scale, 
we then predict that the parameters contribute with at most a factor of $10^{-2}$ to the standard matter effect, decreasing quadratically.

@@

\subsection{DeepCore and PINGU Signals}
Now we repeat our proability analysis but for the DeepCore/PINGU region of \SIrange{5.6}{56}{\GeV}. As we previously saw,
we have rapid oscillations, which means that "indirect" modifications (i.e. $\eet$ will affect the $\Pmm$ channel)
will be more apparent, since all flavors are involved to a greater degree compared with the more stable region above \SI{500}{\GeV}, where 
many oscillations have settled down.

@@

>>>>

Compared to other fermions, the lightness of neutrino masses along with their sparse SM interactions might 
indicate that these particles provide the best starting point for us to probe new physics, in which 
exotic neutrino interactions might occur~\cite{gavela2009}. We call these interactions non-standard interactions (NSI) in order to 
distinguish them from the `standard' interactions of the Standard Model.

Following the approach of the Standard Model that the group generators uniquely determine the gauge bosons, a different gauge
theory will have different interactions than those we presently know. These new interactions can be parametrized as model-independent four-fermion effective operators \cite{salvadoNSI,nsiFarzan}.
Following the discussion in~\cite{tommyNSI}, we see that the NSI parameters $\epsilon$ resulting from six-dimensional operators have the scale
\begin{align}
    \epsilon \propto \frac{m_W^2}{m_{\epsilon}^2} \sim \frac{10^{-2}}{m_\epsilon^2}\,
\end{align} 
in TeV, so the new interactions generated at a mass scale of $m_\epsilon = \SI{1}{TeV}$ will produce parameters in the order of $10^{-2}$, 
two magnitudes below the standard matter effect. Thus, if we assume the new interactions to arise from a higher-energy theory above electroweak scale, 
we then predict that the parameters contribute with at most a factor of $10^{-2}$ to the standard matter effect, decreasing quadratically.

@@

\subsection{DeepCore and PINGU Signals}
Now we repeat our proability analysis but for the DeepCore/PINGU region of \SIrange{5.6}{56}{\GeV}. As we previously saw,
we have rapid oscillations, which means that `indirect' modifications (i.e. $\eet$ will affect the $\Pmm$ channel)
will be more apparent, since all flavors are involved to a greater degree compared with the more stable region above \SI{500}{\GeV}, where 
many oscillations have settled down.

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=0.8\textwidth]{figures/DC_event_pulls.pdf}
        \caption{Plots of the quantity $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$, displaying which 
        reconstructed bins the NSI signal is most clear.}
        \label{fig:event_pulls}
    \end{center}
\end{figure}
%TODO: talk a bit about this. Compare cascades to tracks

### 2021-06-16 thesis/tex/oscillations.tex
update

\subsection{Mass Generation}
In the Standard Model, fermion masses are generated by the Higgs mechanism through Yukawa couplings with the fermions right and left-handed components.
All neutrinos are left-handed and all antineutrinos are right-handed. Thus, neutrinos are massless since they do not undergo the Higgs mechanism.
Additionally, all terms of a Lagrangian that we construct must respect the gauge invariance. This removes the possibility of the neutrino mass to be 
generated at loop level because any such attempt will violate the total lepton number by two units.
In order to keep our theory renormalizable, we have one option left if we want to generate neutrino masses: introducing additional fermions.

@@

\begin{align}\label{eq:avg_H1}
    \avg{H_{CC}} &= \int \dd p_e^3 \bra{e(p_e,h_e)} \times \frac{1}{2} \sum_{h_e} H   f(E_e,T) \ket{e(p_e,h_e)} \nonumber \\
           &= \frac{G_F}{\sqrt{2}} \int \dd p_e^3 \bra{e(p_e,h_e)} \left[ \bar\ne \gamma^\rho\, (1- \gamma^5)\, \ne \right]  f(E_e,T) \times \frac{1}{2} \sum_{h_e} \left[\bar e(x) \,\gamma_\rho\, (1- \gamma^5)\, e(x) \right] \ket{e(p_e,h_e)} \nonumber \\
           &= \frac{G_F}{\sqrt{2}} \bar\ne \gamma^\rho\, (1- \gamma^5)\, \ne \int \dd p_e^3  f(E_e,T) \times \frac{1}{2} \sum_{h_e} \bra{e(p_e,h_e)} \bar e(x) \,\gamma_\rho\, (1- \gamma^5)\, e(x)   \ket{e(p_e,h_e)}\,.
\end{align}
First, calculate the sum using trace technology %TODO: explain this more
\begin{align}\label{eq:helicity_sum}
    \frac{1}{2} \sum_{h_e} \bra{e(p_e,h_e)} \bar e(x) \,\gamma_\rho\, (1- \gamma^5)\, e(x)   \ket{e(p_e,h_e)} &= \frac{1}{4E_e V} \sum_{h_e} \bar{u}_e^{h_e}(p_e) \,\gamma_\rho\, (1- \gamma^5)\, u_e^{h_e}(p_e) \nonumber \\
    &= \frac{1}{4 E_e V} \Trace{\left[ \sum_{h_e} \bar{u}_e^{h_e}(p_e) u_e^{h_e}(p_e) \,\gamma_\rho\, (1- \gamma^5)\, \right]} \nonumber \\
    &= \frac{1}{4 E_e V} \Trace{\left[ (\slashed{p}_e + m_e ) \,\gamma_\rho\, (1- \gamma^5)\, \right]} \nonumber \\
    &= \frac{(p_e)_\rho}{E_e V}\,.
\end{align}
Eq.~\ref{eq:avg_H1} now becomes 
\begin{align}\label{eq:avg_H2}
    \avg{H_{CC}} &= \frac{G_F}{\sqrt{2}E_e V} \bar\ne \, (1- \gamma^5)\, \ne \int \dd p_e^3\, \slashed{p}_e\, f(E_e,T)\,.
\end{align}
Expand the integral, and use the fact that $\vec{p}_e$ is odd:
\begin{align}
    \int \dd p_e^3\, \slashed{p}_e\, f(E_e,T) &= \int \dd p_e^3\, f(E_e,T) (\gamma^0 E_e - \vec{p}_e \cdot \vec{\gamma}) \nonumber \\
                                              &= \int \dd p_e^3\, f(E_e,T) \gamma^0 E_e \nonumber \\
                                              &= \gamma_0 E_e N_e V\,.
\end{align}
Inserting this into Eq.~\ref{eq:avg_H2}, we have
\begin{align}
    \avg{H_{CC}} &= \frac{G_F N_e}{\sqrt{2}} \bar{\nu}_e \, (1- \gamma^5)\, \ne \gamma_0 \nonumber \\
            &= \sqrt{2}G_F N_e \bar{\nu}_{eE} \gamma^0 \nu_{eL}\,,
\end{align}
where the projection operator $(1- \gamma^5)$ in the first line ensures that only the left-hand compontent of the neutrino fields interact. Thus,
\begin{align}\label{eq:V_CC1}
    V_{CC} = \sqrt{2}G_F N_e\,, \quad H_{CC}\ket{\nu_k} = V_{CC}\ket{\nu_k}\,.
\end{align}
Here we see a crucial difference between the eigenvectors between the vacuum Hamiltonian defined in Eq.~\ref{eq:osc_1} and $H_{CC}$, namely
that the CC interactions happen in the flavor basis rather than in the mass basis\footnote{A similar calculation for the NC reveals the same fact:
neutrinos interact in the flavor basis.}. In other words,
neutrinos propagate in their mass eigenstates, but interact in their flavor eigenstate. The mixing of mass eigenstates during propagation
determines if the flavor eigenstate has oscillated or not. The interactions still conserve lepton number, but the kinetic Lagrangian will not.

@@

Using a value of $Y=0.5$ electron per nucleon, we express the matter potential as
\begin{align}
    V_{CC} &= \sqrt{2}G_F N_e = \sqrt{2}G_F Y N_A \,\rho \nonumber \\
           &= \SI{3.8e-23}{\eV\, \gram \cm^{-3}}\,,
\end{align}
a low number due to the smallness of $G_F$. This is plotted in the right panel of Fig.~\ref{fig:potential}
\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/potential.pdf}
    \caption{\emph{Left panel:} The spherically symmetric Earth density according to PREM~\cite{PREM}, as of distance from the core.
    \emph{Right panel:} $V_{CC}$ using the PREM density and 1/2 electrons per nucleon.}\label{fig:potential}
\end{figure}

>>>>

\subsection{Mass Generation}
In the Standard Model, fermion masses are generated by the Higgs mechanism through Yukawa couplings with the fermion's right and left-handed components.
All neutrinos are left-handed and all antineutrinos are right-handed. Thus, neutrinos are massless since they do not undergo the Higgs mechanism.
Additionally, all terms of a Lagrangian that we construct must respect the gauge invariance. This removes the possibility of the neutrino mass to be 
generated at loop level because any such attempt will violate the total lepton number by two units.
In order to keep our theory renormalizable, we have one option left if we want to generate neutrino masses: introducing additional fermions.

@@

\begin{align}\label{eq:avg_H1}
    \avg{H_{CC}} &= \int \dd p_e^3 \bra{e(p_e,h_e)} \times \frac{1}{2} \sum_{h_e} H   f(E_e,T) \ket{e(p_e,h_e)} \nonumber \\
           &= \frac{G_F}{\sqrt{2}} \int \dd p_e^3 \bra{e(p_e,h_e)} \left[ \bar\ne \gamma^\rho\, (1- \gamma^5)\, \ne \right]  f(E_e,T) \times \frac{1}{2} \sum_{h_e} \left[\bar e(x) \,\gamma_\rho\, (1- \gamma^5)\, e(x) \right] \ket{e(p_e,h_e)} \nonumber \\
           &= \frac{G_F}{\sqrt{2}} \bar\ne \gamma^\rho\, (1- \gamma^5)\, \ne \int \dd p_e^3  f(E_e,T) \times \frac{1}{2} \sum_{h_e} \bra{e(p_e,h_e)} \bar e(x) \,\gamma_\rho\, (1- \gamma^5)\, e(x)   \ket{e(p_e,h_e)}\,.
\end{align}
First, calculate the sum using trace technology
\begin{align}\label{eq:helicity_sum}
    \frac{1}{2} \sum_{h_e} \bra{e(p_e,h_e)} \bar e(x) \,\gamma_\rho\, (1- \gamma^5)\, e(x)   \ket{e(p_e,h_e)} &= \frac{1}{4E_e V} \sum_{h_e} \bar{u}_e^{h_e}(p_e) \,\gamma_\rho\, (1- \gamma^5)\, u_e^{h_e}(p_e) \nonumber \\
    &= \frac{1}{4 E_e V} \Trace{\left[ \sum_{h_e} \bar{u}_e^{h_e}(p_e) u_e^{h_e}(p_e) \,\gamma_\rho\, (1- \gamma^5)\, \right]} \nonumber \\
    &= \frac{1}{4 E_e V} \Trace{\left[ (\slashed{p}_e + m_e ) \,\gamma_\rho\, (1- \gamma^5)\, \right]} \nonumber \\
    &= \frac{(p_e)_\rho}{E_e V}\,.
\end{align}
Eq.~\ref{eq:avg_H1} now becomes 
\begin{align}\label{eq:avg_H2}
    \avg{H_{CC}} &= \frac{G_F}{\sqrt{2}E_e V} \bar\ne \, (1- \gamma^5)\, \ne \int \dd p_e^3\, \slashed{p}_e\, f(E_e,T)\,.
\end{align}
Expand the integral, and use the fact that $\vec{p}_e$ is odd:
\begin{align}
    \int \dd p_e^3\, \slashed{p}_e\, f(E_e,T) &= \int \dd p_e^3\, f(E_e,T) (\gamma^0 E_e - \vec{p}_e \cdot \vec{\gamma}) \nonumber \\
                                              &= \int \dd p_e^3\, f(E_e,T) \gamma^0 E_e \nonumber \\
                                              &= \gamma_0 E_e N_e V\,.
\end{align}
Inserting this into Eq.~\ref{eq:avg_H2}, we have
\begin{align}
    \avg{H_{CC}} &= \frac{G_F N_e}{\sqrt{2}} \bar{\nu}_e \, (1- \gamma^5)\, \ne \gamma_0 \nonumber \\
            &= \sqrt{2}G_F N_e \bar{\nu}_{eE} \gamma^0 \nu_{eL}\,,
\end{align}
where the projection operator $(1- \gamma^5)$ in the first line ensures that only the left-hand compontent of the neutrino fields interact. Thus,
\begin{align}\label{eq:V_CC1}
    V_{CC} = \sqrt{2}G_F N_e\,, \quad H_{CC}\ket{\nu_k} = V_{CC}\ket{\nu_k}\,.
\end{align}
Here we see a crucial difference between the eigenvectors between the vacuum Hamiltonian defined in Eq.~\ref{eq:osc_1} and $H_{CC}$, namely
that the CC interactions happen in the flavor basis rather than in the mass basis\footnote{A similar calculation for the NC reveals the same fact:
neutrinos interact in the flavor basis.}. In other words,
neutrinos propagate in their mass eigenstates, but interact in their flavor eigenstate. The mixing of mass eigenstates during propagation
determines if the flavor eigenstate has oscillated or not. The interactions still conserve lepton number, but the kinetic Lagrangian will not.

@@

Using a value of $Y=0.5$ electron per nucleon, we express the matter potential as
\begin{align}
    V_{CC} &= \sqrt{2}G_F N_e = \sqrt{2}G_F Y N_A \,\rho \nonumber \\
           &= \SI{3.8e-23}{\eV \gram \cm^{-3}}\,,
\end{align}
a low number due to the smallness of $G_F$. This is plotted in the right panel of Fig.~\ref{fig:potential}
\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/potential.pdf}
    \caption{\emph{Left panel:} The spherically symmetric Earth density according to PREM~\cite{PREM}, as of distance from the core.
    \emph{Right panel:} $V_{CC}$ using the PREM density and 1/2 electrons per nucleon.}\label{fig:potential}
\end{figure}

### 2021-06-16 thesis/tex/sterileTheory.tex
update

The experiments listed above indicate that the mass-squared difference of the 
sterile neutrino is in the \si{\eV\squared} scale, while the two others
are three and five magnitudes smaller. To remind us of this difference, we write $3+1$, since the sterile state 
proposed is a lot more massive than the active states.

@@

Now, looking at Eqs.~\ref{eq:U4gen_param} and~\ref{eq:dm4gen}, we see that we have introduced four new parameters: $\dm[41]$, $\theta_{14}$
, $\theta_{24}$, $\theta_{34}$.

@@

\begin{figure}
    \centering
    \begin{subfigure}{0.45\textwidth}
        \includegraphics[width=1\linewidth]{figures/sterile_resonance.pdf}
        \caption{$\Pamam$ disappearance}\label{fig:sterile_resonance}
    \end{subfigure}
    \begin{subfigure}{0.45\textwidth}
        \includegraphics[width=1\linewidth]{figures/resonance_shift.pdf}
        \caption{$\Pamam$ disappearance}\label{fig:resonance_shift}
    \end{subfigure}
\end{figure} %TODO: better captions

So for a \si{\eV}-scale sterile neutrino and a non-zero $\theta_{24}$, we expect a \si{\TeV} $\anm$ disappearance.
The resonance dip is affected by the value of $\dm[41]$ and $\theta_{24}$ as shown in Fig.~\ref{fig:resonance_shift}.
We see how the value of $\dm[41]$ shifts the peak, while the mixing angle adjusts its strength. 
%TODO: more? mention that Pmm=1 at TeV for 3+0

>>>>

The experiments listed above indicate that the mass-squared difference of the 
sterile neutrino is in the \si{\eV\squared} scale, while the two others
are three and five magnitudes smaller. To remind us of this hierarchy, we write $3+1$, since the sterile state 
proposed is a lot more massive than the active states.

@@

Now, looking at Eqs.~\ref{eq:U4gen_param} and~\ref{eq:dm4gen}, we see that we have introduced four new parameters: $\dm[41]$, $\theta_{14}$
, $\theta_{24}$, $\theta_{34}$. In our analysis, we reduce the number of parameters by only considering $\dm[41]$ and $\theta_{24}$, which are 
the most important for the $\nm$ oscillations. In fact, one can close the $e$ channel by letting $\theta_{14} = 0$ in the \si{\GeV} to \si{\TeV} range, thus 
using a two-flavor neutrino approximation. This is not done in this work.

@@

\begin{figure}
    \centering
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=1\linewidth]{figures/sterile_resonance.pdf}
        \caption{$\Pamam$ resonant disappearance}\label{fig:sterile_resonance}
    \end{subfigure}
    \begin{subfigure}{0.69\textwidth}
        \includegraphics[width=1\linewidth]{figures/resonance_shift.pdf}
        \caption{Shifting and amplification of the resonance by altering $\dm[41]$ and $\theta_{24}$, respectively.}\label{fig:resonance_shift}
    \end{subfigure}
\end{figure}

So for a \si{\eV}-scale sterile neutrino and a non-zero $\theta_{24}$, we expect a \si{\TeV} $\anm$ disappearance.
The resonance dip is affected by the value of $\dm[41]$ and $\theta_{24}$ as shown in Fig.~\ref{fig:resonance_shift}.
We see how the value of $\dm[41]$ shifts the peak, while the mixing angle adjusts its strength. 
In the $3+0$ scenario, $\Pamam = 1$ in this range. Thus, if we are able to see a $\anm$ disappearance in the \si{\TeV} region,
we are in a good position to argue for a \si{\eV} scale $\dm[41]$.
